Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.

### 2022-07-11 manuscript.Rmd
remove old materials

$$
P(H|E) = \frac{P(H)P(E|H)}{P(E)}
$$

@@

$$
P(E) = P(H)P(E|H) + P(¬H)P(E|¬H)
$$

@@

Both theorethical risk ($P(¬E|¬H)$) and detectability ($P(E|H)$) aggregate uncountable influences, otherwise they could not model the process of evidential support for theories.
To illustrate the concepts we introduced up to here, consider the following  example of a single theory and three experiments that may test it. 
The experiments were created to illustrate how they may differ in their theoretical risk and detectability.
Suppose the primary theory is about the cognitive phenomenon of "insight".
For the purpose of illustration, we define it somewhat hand-wavinlyg as an cognitive abstraction that allows agents to consistently solve a well-defined class of problems.
We pose the hypothesis that the following problem belongs to such class of insight problems:

@@

$$
\mathbb{E}(p(H|E)) = \int \frac{p(H)p(E|H)}{p(H)p(E|H)+p(\neg H) p(E|\neg H) } \; \text{d}\; \mathbb{P}(p(E|\neg H))
$$

@@

$$
f(x)=\begin{cases}
  \frac{1}{\tau} & \mathrm{for}\ 0 \le x \le \tau,\\
  0 & \mathrm{for}\ x<0\ \mathrm{or}\ x>\tau
  \end{cases}
$$

@@

# Old Material

> As far as I am aware, Mayo’s severity argument currently provides one of the few philosophies of science that allows for a coherent conceptual analysis of the value of preregistration.
> Borsboom

We, therefore, propose an interpretation of preregistration that increases the expected epistemic value of studies without necessarily increasing the theoretical risk.

If you are a follow a hypo-deductive rational this is self evident, if you are positivist we hope our

The other two assumptions are later necessary to connect individual decisions that affect epistemic value to the research community as a whole.

Observe, that since $P(E|¬H)$.

Please note, that we adopt a Bayesian rational for the meta scientific process of preregistration but that this does not imply any ties to the methods a researcher uses.

However, we aim to show that preregistration is indispensable for adequately judging a much broader range of studies.
To that end, we first rationalize the appreciation for conformation using Bayesian reasoning.
Equipped with some of the basic tools of Bayesian Philosophy of Science, we can move beyond a simple dichotomy of exploration and confirmation.

To that end we show that the appreciation for the distinction between exploration and confirmatory neatly falls in line with Bayesian reasoning connected through a quantity we call *theorethical risk*.
Than we proceed to show that preregistration impacts the theorethical risk via two pathways.
Only considering the one path rationalizes recommendation to employ preregistration only for confirmatory studies, taking both paths together into consideration leads to a much wider applicability of preregistration.

But what exactly have you been cheated of?
This appreciation of confirmatory results falls neatly in line with Bayesian reasoning (XXX).
<!-- We leverage this reasoning to outline two paradigms that rationalize the use of preregistration. -->
In order to apply Bayesian reasoning we have to cast conformation and exploration in terms of probability.
Consider the hypothesis that people who regularly engage in fitness activities are healthier.
A confirmatory story might operationalize regularly as weekly, fitness activity as light jogging and healthy as blood cholesterol level.
A more exploratory study might consider several definitions of "regularly", "fitness activities", and "healthy", e.g. "monthly", "weekly", "daily", or "marathon", "light jogging", "intense cleaning", or "cholesterol", "self rated health", "ability to play tag".
Both studies have some probability to turn up with positive results even in a world where fitness activities do not lead to healthier people, because there are other plausible explanations for the results.
However, for the second study the number of explanations that could lead to evidence in favor, aside from the theory is much greater.
Suppose a researcher actually conducted the second study but is reporting it as if they conducted the first study.
You would feel cheated.
But what exactly have you been cheated of?
In fact, you have been cheated out of the information about all the possibilities that could have brought about this evidence.
We summarize these possibilities as a probability.
Therefore, we argue that confirmatory studies have a low probability of observing evidence in favor of their theory if we assume the theory to be false.
Exploratory studies on the other hand are characterized by a higher probability that they find evidence, even when their hypothesis turns out to be wrong.

We connect the question of exploration vs confirmation to Bayesian reasoning via a quantity we call theoretical risk (we borrow the term from Meehl).

Theoretical risk is the inverse of the probability that one observes evidence in favor of a theory when we assume that the theory does not hold.
We assume that confirmatory studies take a high theoretical risk, compared to exploratory studies which take a low theoretical risk.
The tighter your definition of "evidence in favor" is, the more the probability of observing the evidence is tied to your theory holding.

<!-- Caroline does not see the connection to type I error rate/theoretical risk -->

Let us formalize the judgment about a hypothesis (H) given some evidence (E) as the conditional probability $p(H|E)$.

Using Bayes' rule we arrive at:

$$
p(H|E) = \frac{p(H)p(E|H)}{p(H)p(E|H) + p(¬H)p(E|¬H)}
$$

Note the connection to null hypothesis testing.
If H represents the alternative Hypothesis and ¬H represents the Null hypothesis, then $p(E|H)$ is the power, while $p(E|¬H)$ represents the type I/alpha error.
However, generally a theory is richer in content, than equating it with the alternaive hypothesis in Null hypothesis testing.

<!-- continue to justify the importance of theorethical risk -->

The first paradigm is embraced by the open science community and conceptualizes preregistration as a tool to reduce the type I error rate (XXXX).
As we will later show, reducing the type I error rate is indeed a powerful leaver to gather persuasive evidence.
In practice, however, researchers can not reduce the type I error rate at will.
They face three limiting factors.
First, psychological theories tend to be vague.
Some decisions are just not derivable from theory, which makes it extraordinarily difficult to maintain a type I error rate near zero.
Second, researchers are constrained by resource limitations.
One way to counter the first problem is conducting exploratory pilot studies, but running extensive pilot studies is expensive.
Doubling the resources that go into investigating a research question is often just not tenable and mostly beyond the individual researcher to decide.
Third, by reducing the type I error rate the researcher directly increases the likelihood that their study will result in disappointing null results.
A study that fails to deliver results can be crippling to the researchers career.
Depending on individual circumstances, e.g. the theory in question, resources allocated to a study, the career stage of the researcher, a researcher may simply feel unable to reduce the type I error rate to the threshold necessary for a confirmatory study.
Constrained by these factors researchers tasked with writing a preregistration are incentiviced to either give up or to muddle through by being intentionally vague. The latter practice leads to preregistrations stating unspecific assumptions such as "we expect some of the Xs to correlate with some of the Ys".

Our conclusions can be applied to bayesian and frequentist methods alike.
Of course, frequentist methods are traditionally viewed from a Popperian (or hypo-deductive) philosophy of science, but our conclusions are compatible with this view and notable later advancements like the error statistical view.
However, error statistical view marries a statistical philosophy with specific statistical models and procedures.
A marriage which leaves us with richer set of assumptions than necessary for the points we want to make about the value of preregistration.

>>>>

\begin{equation}
P(H|E) = \frac{P(H)P(E|H)}{P(E)} (\#eq:bayes)
\end{equation}

@@

\begin{equation}
P(E) = P(H)P(E|H) + P(¬H)P(E|¬H) (\#eq:total-expectedness)
\end{equation}

@@

Both theorethical risk ($P(¬E|¬H)$) and detectability ($P(E|H)$) aggregate uncountable influences, otherwise they could not model the process of evidential support for theories.
To illustrate the concepts we introduced up to here, consider the following  example of a single theory and three experiments that may test it. 
The experiments were created to illustrate how they may differ in their theoretical risk and detectability.
Suppose the primary theory is about the cognitive phenomenon of "insight".
For the purpose of illustration, we define it somewhat hand-wavingly as an cognitive abstraction that allows agents to consistently solve a well-defined class of problems.
We pose the hypothesis that the following problem belongs to such class of insight problems:

@@

\begin{equation}
\mathbb{E}(p(H|E)) = \int \frac{p(H)p(E|H)}{p(H)p(E|H)+p(\neg H) p(E|\neg H) } \; \text{d}\; \mathbb{P}(p(E|\neg H)) (\#eq:integral)
\end{equation}

@@

\begin{equation}
f(x)=\begin{cases}
  \frac{1}{\tau} & \mathrm{for}\ 0 \le x \le \tau,\\
  0 & \mathrm{for}\ x<0\ \mathrm{or}\ x>\tau
  \end{cases}
\end{equation}

@@
